---
title: "Rmarkdown practice"
author: "H2 Data Science"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      root.dir = rprojroot::find_rstudio_root_file())
```

Congratulations! You've successfully downloaded a repository from GitHub!

Now it's time to practice some R Markdown. R Markdown is an extension of Markdown (a text-based way to format documents) that can embed R code, R output, and R plots. An R Markdown document can be rendered into a clean report in, e.g., PDF or HTML format.

For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

You can toggle back and forth between the Source (the plain text) and the Visual (formatted) views at the top of the file in RStudio.

# Adding R chunks

R code can be inserted in 'code chunks'. Move your cursor to the space below, then click the 'insert code chunk' icon at the top right of this Rmd: it is a green square with a 'c' and a '+'. Select 'R' if you get the option.

In your new code chunk, write code to make a plot using the `iris` dataset.

**Add a code chunk on the next line:**

When working in a .Rmd, you can run lines of code with ctrl+enter (Mac: command+enter). This runs the lines individually and may re-set plotting parameters (e.g., `par(mfrow=...))`. To run the whole chunk at once, click the green arrow on the right side of the chunk, or use ctrl+shift+enter.

# R Markdown and directories

Rmd files can be fiddly with working directories. For our purposes, it is best to always use the project directory as the working directory.

On the top toolbar, click *Tools \> Global options...* and then click on the *R Markdown* tab on the left. Set 'Evaluate chunks in directory:' to 'Project'.

Click OK.

Create a new R chunk below and try loading the file `data/trees.csv`. If it loads without an error, then the working directory is set to the project directory.

# Knitting

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

Try clicking **Knit** at the top of this document. If you get an error, you may need to first run this in R:

```{r}
# tinytex::install_tinytex()
```

If you get an error related to the `data/trees.csv` location, then the **Knit** button is using the wrong working directory (i.e., it is using the `code` folder where this .Rmd is located). If your RStudio version has a drop down menu on the **Knit** button, click it, select 'Knit Directory', and set it to 'Project Directory'. If you do not have that option, you can add `knitr::opts_chunk$set(root.dir = rprojroot::find_rstudio_root_file())` to the first `setup` code chunk at the top of the Rmd file.

Once you have successfully knitted your document, you will see `code/example.pdf` in your files.
